Summary for August, 2021

### 2021-10-31 summaries/21-11/21-11.tex
Updated sensitivity model

\begin{document}
\title{New math derivation}
\author{Jack Dinsmore, Julien de Wit}
\email{jtdinsmo@mit.edu}
\date{\today}
\affiliation{MIT Department of Physics}

@@

\begin{abstract}
    The purpose of this summary is to derive the math governing the way an asteroid reacts to a non-point source. It was motivated by the discovery that, under the assumption of a gravitating point source and a small asteroid, the tidal torque can be written in terms of components of the moment of inertia, which means inference is limited to the six components of the inertia matrix. By violating these assumptions and considering other effects (oblateness, etc.), we can break the degeneracy.
\end{abstract}

@@

The differential force is $d\bm F = -dm \nabla_{\bm R} V(\bm R)$, where
$$V(\bm R) = -\int d^3 r' \frac{\rho_M(\bm r')}{|\bm r' - \bm R|}$$
is the gravitational potential field of the planet, and $\bm R = \bm r - \bm D$ points from $dm$ to the center of mass of the planet. Also, we have $\rho_M$ as the mass distribution of the planet. Expanding this potential field using equation \ref{eqn:expansion}, we get
$$V(\bm R) = -\sum_{lm}S^*_{lm}(\bm R)\int d^3 r' \rho_M(\bm r') R_{lm}(\bm r').$$
We introduce the useful symbol
\begin{equation}
\J_{lm} = \int d^3 r \rho_M(\bm r) R_{lm} (\bm r)
\label{eqn:jlm-def}
\end{equation}
and we have
\begin{equation}
V(\bm R) = -\sum_{lm} S^*_{lm} (\bm R) \J_{lm}.
\label{eqn:potential}
\end{equation}
Incidentally, since $R_{00} = 1$, we have that $\J_{00}$ is the total mass of the asteroid and $\mu = G \J_{00}$, which can be used to simplify equation \ref{eqn:linear-eom}.

The total torque on the asteroid can therefore be written as
\begin{equation}
\bm \tau = \int d^3 r \rho(\bm r) \bm r \times \nabla_{\bm R} \sum_{lm} S^*_{lm}(\bm R) \J_{lm}.
\label{eqn:first-torque}
\end{equation}
Here, $\rho(\bm r)$ is now the mass density of the asteroid. Since $\nabla_{\bm R} = \nabla_{\bm r}$, we will drop the subscript henceforth.

@@

We expand equation \ref{eqn:first-torque} using equation \ref{eqn:trans}:
\begin{equation*}
\begin{aligned}
\bm \tau = &\sum_{lm}(-1)^l\J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) \bm r \times \nabla R^*_{l'm'}(\bm r).
\end{aligned}
\end{equation*}
Now we can move straight on to substitution equation \ref{eqn:grad}:
\begin{equation*}
\begin{aligned}
\bm \tau = &-\frac{1}{2}\sum_{lm}(-1)^l\J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)].
\end{aligned}
\end{equation*}
By defining the new varialble
\begin{equation}
\K_{lm} = \int d^3 r \rho(\bm r) R_{lm} (\bm r),
\label{eqn:klm-def}
\end{equation}
we can simplify this to
\begin{equation*}
\begin{aligned}
\bm \tau = &-\frac{1}{2}\sum_{lm}(-1)^l\J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) \\
&\bigg[(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}\\
&+(i\x-\y)(l'+m'+1)\K^*_{l',m'+1}+2im'\z \K^*_{l'm'}\bigg].
\label{eqn:complex-torque}
\end{aligned}
\end{equation*}

\section{Proof of real-ness}
This equation had better be real. To prove it, let us combine the $\pm m'$ terms, leaving out the $m'=0$ term and using the subscript o (off-diagonal) to indicate its absence:
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm}(-1)^l\J_{lm}\sum_{l'm'_+}\Bigg\{S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)]\\
&S_{l+l',m-m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'+m'+1)R^*_{l',-m'-1}(\bm r)\\
&+(i\x-\y)(l'-m'+1)R^*_{l',-m'+1}(\bm r)-2im'\z R^*_{l',-m'}(\bm r)]\Bigg\}.
\end{aligned}
\end{equation*}
Using the symmetry rules for complex conjugates,
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm}(-1)^l\J_{lm}\sum_{l'm'_+}\Bigg\{S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)]\\
&S_{l+l',m-m'}(\bm D) \\
&(-1)^{m'}\int d^3 r \rho(\bm r) [-(i\x + \y)(l'+m'+1)R_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)R_{l',m'-1}(\bm r)-2im'\z R_{l',m'}(\bm r)]\Bigg\}.
\end{aligned}
\end{equation*}
Now we expand for $\pm m$ too:
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm_+}(-1)^l \Bigg\{\J_{lm}\sum_{l'm'_+} \Bigg[S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)]\\
&+S_{l+l',m-m'}(\bm D) \\
&(-1)^{m'}\int d^3 r \rho(\bm r) [-(i\x + \y)(l'+m'+1)R_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)R_{l',m'-1}(\bm r)-2im'\z R_{l',m'}(\bm r)]\Bigg]\\
&+ \J_{l,-m}\sum_{l'm'_+} \Bigg[S_{l+l',-m+m'}(\bm D) \\
&+\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)]\\
&+S_{l+l',-m-m'}(\bm D) \\
&(-1)^{m'}\int d^3 r \rho(\bm r) [-(i\x + \y)(l'+m'+1)R_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)R_{l',m'-1}(\bm r)-2im'\z R_{l',m'}(\bm r)]\Bigg]\Bigg\}.
\end{aligned}
\end{equation*}
Let us introduce a new notation:
\begin{equation}
\K_{lm} = \int d^3 r \rho(\bm r) R_{lm} (\bm r)
\label{eqn:klm-def}
\end{equation}
which are the expansion coefficients of the potential of the asteroid. Then the above becomes
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm_+}(-1)^l \Bigg\{\J_{lm}\sum_{l'm'_+} \Bigg[S_{l+l',m+m'}(\bm D) \\
&\bigg[(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)K^*_{l',m'+1}(\bm r)+2im'\z \K^*_{l'm'}(\bm r)\bigg]\\
&+S_{l+l',m-m'}(\bm D) \\
&(-1)^{m'}\bigg[-(i\x + \y)(l'+m'+1)\K_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)\K_{l',m'-1}(\bm r)-2im'\z \K_{l',m'}(\bm r)\bigg]\Bigg]\\
&+ (-1)^m (-1)^{m+m'}\J_{l,m}^*\sum_{l'm'_+} \Bigg[S^*_{l+l',m-m'}(\bm D) \\
&\bigg[(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)\K^*_{l',m'+1}(\bm r)+2im'\z \K^*_{l'm'}(\bm r)\bigg]\\
&+S^*_{l+l',m+m'}(\bm D) \\
&(-1)^{m'}\bigg[-(i\x + \y)(l'+m'+1)\K_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)\K_{l',m'-1}(\bm r)-2im'\z \K_{l',m'}(\bm r)\bigg]\Bigg]\Bigg\}.
\end{aligned}
\end{equation*}
which can be simplified as
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm_+}\sum_{l'm'_+}(-1)^l \Bigg\{\J_{lm} \Bigg[S_{l+l',m+m'}(\bm D) \\
&\bigg[(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)K^*_{l',m'+1}(\bm r)+2im'\z \K^*_{l'm'}(\bm r)\bigg]\\
&+S_{l+l',m-m'}(\bm D) \\
&(-1)^{m'}\bigg[-(i\x + \y)(l'+m'+1)\K_{l',m'+1}(\bm r)\\
&-(i\x-\y)(l'-m'+1)\K_{l',m'-1}(\bm r)-2im'\z \K_{l',m'}(\bm r)\bigg]\Bigg]\\
&+\J_{lm}^*\Bigg[S^*_{l+l',m+m'}(\bm D) \\
&\bigg[-(i\x-\y)(l'-m'+1)\K_{l',m'-1}(\bm r)\\
&-(i\x + \y)(l'+m'+1)\K_{l',m'+1}(\bm r)-2im'\z \K_{l',m'}(\bm r)\bigg]\Bigg]\\
&+(-1)^{m'}S^*_{l+l',m-m'}(\bm D) \\
&\bigg[(i\x-\y)(l'+m'+1)\K^*_{l',m'+1}(\bm r)\\
&+(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}(\bm r)+2im'\z \K^*_{l'm'}(\bm r)\bigg]\Bigg\}
\end{aligned}
\end{equation*}
or
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&(l'-m'+1)\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)(i\x + \y)\K^*_{l',m'-1}\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)(i\x-\y)\K_{l',m'-1}\\
&-\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l',m'-1}(i\x-\y)\\
&+(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)(i\x + \y)\K^*_{l',m'-1}\bigg)\\
&+(l'+m'+1)\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)(i\x-\y)\K^*_{l',m'+1}\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)(i\x + \y)\K_{l',m'+1}\\
&-\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l',m'+1}(i\x + \y)\\
&+(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)(i\x-\y)\K^*_{l',m'+1}\bigg)\\
&+2im'\z\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)\K^*_{l'm'}\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)\K_{l'm'}\\
&-\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l'm'}\\
&+(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)\K^*_{l'm'}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}
This can be written as follows
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\frac{1}{2}\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&(l'-m'+1)\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)(i\x + \y)\K^*_{l',m'-1}\\
&+\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l',m'-1}(-i\x+\y)\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)(i\x-\y)\K_{l',m'-1}\\
&-(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)(-i\x - \y)\K^*_{l',m'-1}\bigg)\\
&+(l'+m'+1)\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)(i\x-\y)\K^*_{l',m'+1}\\
&+\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l',m'+1}(-i\x - \y)\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)(i\x + \y)\K_{l',m'+1}\\
&-(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)(-i\x+\y)\K^*_{l',m'+1}\bigg)\\
&+2im'\z\bigg(\J_{lm}S_{l+l',m+m'}(\bm D)\K^*_{l'm'}\\
&-\J_{lm}^*S^*_{l+l',m+m'}(\bm D)\K_{l'm'}\\
&-(-1)^{m'}\J_{lm}S_{l+l',m-m'}(\bm D)\K_{l'm'}\\
&+(-1)^{m'}\J_{lm}^*S^*_{l+l',m-m'}(\bm D)\K^*_{l'm'}\bigg)\Bigg\},
\end{aligned}
\end{equation*}
or
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&+(l'-m'+1)\bigg(\Re\brackets{\J_{lm}S_{l+l',m+m'}(\bm D)(-i\x+\y)\K^*_{l',m'-1}}\\
&-(-1)^{m'}\Re\brackets{\J_{lm}S_{l+l',m-m'}(\bm D)(i\x-\y)\K_{l',m'-1}}\bigg)\\
&+(l'+m'+1)\bigg(\Re\brackets{\J_{lm}S_{l+l',m+m'}(\bm D)(i\x-\y)\K^*_{l',m'+1}}\\
&-(-1)^{m'}\Re\brackets{\J_{lm}S_{l+l',m-m'}(\bm D)(i\x + \y)\K_{l',m'+1}}\bigg)\\
&-2m'\z\bigg(\Im\brackets{\J_{lm}S_{l+l',m+m'}(\bm D)\K^*_{l'm'}}\\
&-(-1)^{m'}\Im\brackets{\J_{lm}S_{l+l',m-m'}(\bm D)\K_{l'm'}}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}
Define
$$\J_{lml'm'}^\pm = \J_{lm} S_{l+l',m\pm m'}(\bm D)(\pm 1)^{m'+1}.$$
Then
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&+(l'-m'+1)\bigg(\Re\brackets{\J_{lml'm'}^+(-i\x+\y)\K^*_{l',m'-1}}\\
&+\Re\brackets{\J_{lml'm'}^-(i\x-\y)\K_{l',m'-1}}\bigg)\\
&+(l'+m'+1)\bigg(\Re\brackets{\J_{lml'm'}^+(i\x-\y)\K^*_{l',m'+1}}\\
&+\Re\brackets{\J_{lml'm'}^-(i\x + \y)\K_{l',m'+1}}\bigg)\\
&-2m'\z\bigg(\Im\brackets{\J_{lml'm'}^+\K^*_{l'm'}}+\Im\brackets{\J_{lml'm'}^-\K_{l'm'}}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}
Further defining
\begin{equation}
\begin{aligned}
\chi_{lml'm'}^{\pm,s}& = \J_{lml'm'}^\pm \K^*_{l',m'+s}\\
&= (\pm 1)^{m'+1}S_{l+l',m\pm m'}(\bm D)\J_{lm} \K^*_{l',m'+s},
\end{aligned}
\label{eqn:chi}
\end{equation}
we get
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&+(l'-m'+1)\bigg(\Re\brackets{\chi_{lml'm'}^{+-}(i\x+\y)}\\
&+\Re\brackets{\chi_{lml'm'}^{--}(i\x-\y)}\bigg)\\
&+(l'+m'+1)\bigg(\Re\brackets{\chi_{lml'm'}^{++}(i\x-\y)}\\
&+\Re\brackets{\chi_{lml'm'}^{-+}(i\x + \y)}\bigg)\\
&-2m'\z\bigg(\Im\brackets{\chi_{lml'm'}^{+0}}+\Im\brackets{\chi_{lml'm'}^{-0}}\bigg)\Bigg\}
\end{aligned}
\end{equation*}
\comment{(Here I realized I made a mistake earlier; I took the complex conjugate of the top line and forgot to conjugate the vector term, so I do it here)}, or
\begin{equation*}
\begin{aligned}
\bm \tau_o = &-\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&+(l'-m'+1)\bigg(\Re\chi_{lml'm'}^{+-}\y - \Im\chi_{lml'm'}^{+-}\x\\
&-\Re\chi_{lml'm'}^{--}\y - \Im\chi_{lml'm'}^{--}\x \bigg)\\
&+(l'+m'+1)\bigg(-\Re\chi_{lml'm'}^{++}\y - \Im\chi_{lml'm'}^{++}\x\\
&+\Re\chi_{lml'm'}^{-+}\y - \Im\chi_{lml'm'}^{-+}\x\bigg)\\
&-2m'\z\bigg(\Im\chi_{lml'm'}^{+0}+\Im\chi_{lml'm'}^{-0}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}
We can write this as
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&+(l'-m'+1)\bigg(\Im\chi_{lml'm'}^{+-}\x - \Re\chi_{lml'm'}^{+-}\y\\
&+\Re\chi_{lml'm'}^{--}\y + \Im\chi_{lml'm'}^{--}\x \bigg)\\
&+(l'+m'+1)\bigg(\Re\chi_{lml'm'}^{++}\y + \Im\chi_{lml'm'}^{++}\x\\
&\Im\chi_{lml'm'}^{-+}\x - \Re\chi_{lml'm'}^{-+}\y\bigg)\\
&+2m'\z\bigg(\Im\chi_{lml'm'}^{+0}+\Im\chi_{lml'm'}^{-0}\bigg)\Bigg\},
\end{aligned}
\end{equation*}
or, joining vectors,
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&\x \bigg((l'-m'+1)\Im\chi_{lml'm'}^{+-} + (l'-m'+1)\Im\chi_{lml'm'}^{--}\\
&+(l'+m'+1)\Im\chi_{lml'm'}^{++} + (l'+m'+1)\Im\chi_{lml'm'}^{-+}\bigg)\\
&+\y \bigg(-(l'-m'+1)\Re\chi_{lml'm'}^{+-} + (l'-m'+1)\Re\chi_{lml'm'}^{--}\\
&+(l'+m'+1)\Re\chi_{lml'm'}^{++} - (l'+m'+1)\Re\chi_{lml'm'}^{-+}\bigg)\\
&+2m'\z\bigg(\Im\chi_{lml'm'}^{+0}+\Im\chi_{lml'm'}^{-0}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}
Finally,
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&\x \Im\bigg((l'+m'+1)\chi_{lml'm'}^{++}+ (l'-m'+1)\chi_{lml'm'}^{--}\\
&+ (l'-m'+1)\chi_{lml'm'}^{+-} + (l'+m'+1)\chi_{lml'm'}^{-+}\bigg)\\
&+\y \Re\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--} \\
&- (l'-m'+1)\chi_{lml'm'}^{+-} - (l'+m'+1)\chi_{lml'm'}^{-+}\bigg)\\
&+2m'\z\Im\bigg(\chi_{lml'm'}^{+0}+\chi_{lml'm'}^{-0}\bigg)\Bigg\}.
\end{aligned}
\end{equation*}

Note that
\begin{equation}
\begin{aligned}
\chi_{l,-m,l',m'}^{\pm,s} = &(\pm 1)^{m'+1}S_{l+l',-m\pm m'}(\bm D)\J_{l,-m} \K^*_{l',m'+s}\\
=&(-1)^s(\pm 1)^{m'+1}S^*_{l+l',m\mp m'}(\bm D)\J^*_{lm} \K_{l',-m'-s}\\
=&(-1)^s{\chi_{l,m,l',-m'}^{\pm,-s}}^*.
\end{aligned}
\label{eqn:chi}
\end{equation}
and
\begin{equation}
\begin{aligned}
\chi_{l,-m,l',-m'}^{\pm,s} = &(\pm 1)^{-m'+1}S_{l+l',-m\mp m'}(\bm D)\J_{l,-m} \K^*_{l',-m'+s}\\
=&(-1)^s(\pm 1)^{m'+1}S^*_{l+l',m\pm m'}(\bm D)\J^*_{lm} \K_{l',m'-s}\\
=&(-1)^s{\chi_{lml'm'}^{\pm,-s}}^*
\end{aligned}
\label{eqn:chi}
\end{equation}
so
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&\x \Im\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\\
&+ (l'-m'+1)\chi_{l,-m,l',-m'}^{++} + (l'+m'+1)\chi_{l,-m,l',-m'}^{--}\bigg)\\
&+\y \Re\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--} \\
&+ (l'-m'+1)\chi_{l,-m,l',-m'}^{++} + (l'+m'+1)\chi_{l,-m,l',-m'}^{--}\bigg)\\
&+2m'\z\Im\bigg(\chi_{lml'm'}^{+0}+\chi_{lml'm'}^{-0}\bigg)\Bigg\}
\end{aligned}
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\sum_{lm_+}\sum_{l'm'_+}(-1)^l\Bigg\{\\
&\x \Im\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+\y \Re\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+m'\z\Im\bigg(\chi_{lml'm'}^{+0}+\chi_{lml'm'}^{-0}\bigg)\Bigg\}\\
&+\sum_{lm_-}\sum_{l'm'_-}(-1)^l\Bigg\{\\
&\x \Im\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+\y \Re\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+m'\z\Im\bigg(\chi_{lml'm'}^{+0}+\chi_{lml'm'}^{-0}\bigg)\Bigg\}
\end{aligned}
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm \tau_o = &\parens{\sum_{lm_+}\sum_{l'm'_+}+\sum_{lm_-}\sum_{l'm'_-}}(-1)^l\Bigg\{\\
&\x \Im\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+\y \Re\bigg((l'+m'+1)\chi_{lml'm'}^{++} + (l'-m'+1)\chi_{lml'm'}^{--}\bigg)\\
&+m'\z\Im\bigg(\chi_{lml'm'}^{+0}+\chi_{lml'm'}^{-0}\bigg)\Bigg\}
\end{aligned}
\end{equation*}

\clearpage

>>>>

\begin{document}
\title{Summary for August, 2021}
\author{Jack Dinsmore}
\email{jtdinsmo@mit.edu}
\date{\today}
\affiliation{MIT Department of Physics}

@@

\begin{abstract}
    The purpose of this summary is to derive the math governing the way an asteroid reacts to a non-point source. It was motivated by the discovery that, under the assumption of a gravitating point source and a small asteroid, the tidal torque can be written in terms of components of the moment of inertia, which means inference is limited to the six components of the inertia matrix. By violating these assumptions and considering other effects (oblateness, etc.), we can break the degeneracy. This month I correct some old mistakes.
\end{abstract}

@@

The differential force is $d\bm F = -dm \nabla_{\bm R} V(\bm R)$, where
$$V(\bm R) = -\int d^3 r' \frac{\rho_M(\bm r')}{|\bm r' - \bm R|}$$
is the gravitational potential field of the planet, and $\bm R = \bm r - \bm D$ points from $dm$ to the center of mass of the planet. Also, we have $\rho_M$ as the mass distribution of the planet. Expanding this potential field using equation \ref{eqn:expansion}, we get
$$V(\bm R) = -\sum_{lm}S^*_{lm}(\bm R)\int d^3 r' \rho_M(\bm r') R_{lm}(\bm r').$$
We introduce the useful symbol
\begin{equation}
\J_{lm} = \frac{1}{M_M}\int d^3 r \rho_M(\bm r) a_M^{-l} R_{lm} (\bm r)
\label{eqn:jlm-def}
\end{equation}
where $M_M$ is the mass of the Earth and $a_M$ is its radius. Then we have
\begin{equation}
V(\bm R) = -M_M\sum_{lm} a_M^{l} S^*_{lm} (\bm R) \J_{lm}.
\label{eqn:potential}
\end{equation}
Incidentally, since $R_{00} = 1$, we have that $\J_{00}=1$. Choosing coordinates in which the Earth's center of mass is at the origin, we have $\J_{1m}=0$ as well.

The total torque on the asteroid can therefore be written as
\begin{equation}
\bm \tau = GM_M \int d^3 r \rho(\bm r) \bm r \times \nabla_{\bm R} \sum_{lm} a_M^l S^*_{lm}(\bm R) \J_{lm}.
\label{eqn:first-torque}
\end{equation}
Here, $\rho(\bm r)$ is now the mass density of the asteroid. Since $\nabla_{\bm R} = \nabla_{\bm r}$, we will drop the subscript henceforth.

@@

We expand equation \ref{eqn:first-torque} using equation \ref{eqn:trans}:
\begin{equation*}
\begin{aligned}
\bm \tau = &GM_M \sum_{lm}(-1)^l a_M^l \J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) \bm r \times \nabla R^*_{l'm'}(\bm r).
\end{aligned}
\end{equation*}
Now we can move straight on to substitution equation \ref{eqn:grad}:
\begin{equation*}
\begin{aligned}
\bm \tau = &-\frac{GM_M}{2}\sum_{lm}(-1)^l a_M^l \J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) \\
&\int d^3 r \rho(\bm r) [(i\x + \y)(l'-m'+1)R^*_{l',m'-1}(\bm r)\\
&+(i\x-\y)(l'+m'+1)R^*_{l',m'+1}(\bm r)+2im'\z R^*_{l'm'}(\bm r)].
\end{aligned}
\end{equation*}
We define the new varialble
\begin{equation}
\K_{lm} = m\int d^3 a_m^{-l}r \rho(\bm r) R_{lm} (\bm r),
\label{eqn:klm-def}
\end{equation}
where $M_m$ is the mass of the asteroid and $a_m$ is a constant with units of length which we define later. Think of it as something like the radius of the asteroid. Note the convenient fact that
\begin{equation}
\begin{aligned}
\K_{lm}^* = &M_m\int d^3 a_m^{-l}r \rho(\bm r) R_{lm}^* (\bm r) \\
= &(-1)^mM_m\int d^3 a_m^{-l}r \rho(\bm r) R_{l,-m} (\bm r) = (-1)^mK_{l,-m}.\\
\label{eqn:klm-def}
\end{aligned}
\end{equation}

We can simplify the torque to
\begin{equation}
\begin{aligned}
\frac{\bm \tau}{M_m} = &-\frac{GM_M}{2}\sum_{lm}(-1)^l \J_{lm}\sum_{l'm'}S_{l+l',m+m'}(\bm D) a_M^l a_m^{l'}\\
&\bigg[(i\x + \y)(l'-m'+1)\K^*_{l',m'-1}\\
&+(i\x-\y)(l'+m'+1)\K^*_{l',m'+1}+2im'\z \K^*_{l'm'}\bigg].
\label{eqn:complex-torque}
\end{aligned}
\end{equation}
which evidently obeys unit analysis, since $\J_{lm}$ and $\K_{lm}$ are unit-less. This equation is also real, but the proof is nontrivial. We will instead investigate the outcomes of this equation under the simplifying assumptions that $l=m=0$ and $l'=2$. But first, we discuss the torque in relation to the equations of motion.

\section{Moment of inertia}
The moment of inertia matrix of an object is written as
$$I_{xx}=\int d^3 r \rho(\bm r) (y^2 + z^2), \indent \dots$$
$$I_{xy}=-\int d^3 r \rho(\bm r) xy, \indent \dots.$$
Note that
$$x^2=r^2\brackets{-\frac{1}{3}(Y_{20}-1)+2Y_{2,-2}+\frac{1}{12}Y_{22}}$$
$$y^2=r^2\brackets{-\frac{1}{3}(Y_{20}-1)-2Y_{2,-2}-\frac{1}{12}Y_{22}}$$
$$z^2=r^2\brackets{\frac{2}{3}Y_{20}+\frac{1}{3}}$$
$$xz=r^2\brackets{\frac{Y_{21}}{6}-Y_{2,-1}}$$
$$yz=r^2\brackets{\frac{Y_{21}}{6i}+\frac{Y_{2,-1}}{i}}$$
$$xy=r^2\brackets{\frac{Y_{22}}{12i}-2\frac{Y_{2,-2}}{i}}.$$
Then we can write the above in terms of $R_{2m}$:
$$I_{xx}=\int d^3r \rho(\bm r)-\frac{1}{3}(2R_{20}-r^2)+2R_{2,-2}+2R_{22}$$
$$I_{yy}=\int d^3r \rho(\bm r)-\frac{1}{3}(2R_{20}-r^2)-2R_{2,-2}-2R_{22}$$
$$I_{zz}=\int d^3r \rho(\bm r)\frac{4}{3}R_{20}+\frac{r^2}{3}$$
$$I_{xz}=-\int d^3r \rho(\bm r)R_{21}-R_{2,-1}$$
$$I_{yz}=-\int d^3r \rho(\bm r)\frac{R_{21}+R_{2,-1}}{i}$$
$$I_{xy}=-\int d^3r \rho(\bm r)2\frac{R_{22}-R_{2,-2}}{i}$$
and therefore
$$I_{xx}=\parens{\frac{2}{3}\K_{20}-2\K_{2,-2}-2\K_{22}+\frac{2}{3}}M_ma_m^2$$
$$I_{yy}=\parens{\frac{2}{3}\K_{20}+2\K_{2,-2}+2\K_{22}+\frac{2}{3}}M_ma_m^2$$
$$I_{zz}=\parens{-\frac{4}{3}\K_{20} + \frac{2}{3}}M_ma_m^2$$
$$I_{xz}=\parens{\K_{21}-\K_{2,-1}}M_ma_m^2$$
$$I_{yz}=\parens{\frac{\K_{21}+\K_{2,-1}}{i}}M_ma_m^2$$
$$I_{xy}=\parens{-2\frac{\K_{22}-\K_{2,-2}}{i}}M_ma_m^2.$$
Here, we have judiciously defined $a_m$ so that the above equations are true. That definition is
\begin{equation}
a_m^2 = \brackets{\int d^3 r \rho(\bm r) r^2}\brackets{\int d^3 r \rho(\bm r)}^{-1}
\label{eqn:am}
\end{equation}
and it is something like the radius of the asteroid. If the asteroid were a sphere of radius $R$, we would have $a_m = R\sqrt{3/5}$

Thus, the moment of inertia is determined entirely by $\K_{lm}$

\section{Computing torque}
Note that the $S$ present in equation \ref{eqn:complex-torque} forces $l$ and $l'$ to be low in the dominant terms. However, $l'=0$ contributes nothing to torque, nor does $l'=1$ due to the fact that $\K_{1m}=0$, which is caused by $\K_{lm}$ being centered on the center of mass of the asteroid. Therefore, $l'=2$ is the minimum value, and the dominant torque term is proportional to $D^{-3}$.
